Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-29 paper/synthesis.tex
Pass over branching processes

@@

To extend $M$ to consider the new possible constructions, we need to calculate
$M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
and $j$, respectively.
%
Fortunately, it is possible to follow the same pattern as we did on our original
work, this time considering the amount of recursive pattern variables if it the
case of a pattern matching value being a father individual, and the amount of
recursive arguments if we deal with a abstract interface function call as a
father individual.
%
For instance, the average amount of function calls to |div| that we will produce
provided that we generated a value matching the second pattern of |simplify| as a
father results $2\cdot p_{div}$.
%
Similarly, the average amount of values satisfying the first pattern matching of
|simplify| provided that we generated a |Tag| constructor as a father results
$1 \cdot p_{simplify\# 1}$.

Furthermore, we found that the prediction process can be factored in terms of
two vectors $\beta$ and $\mathcal{P}$, such that $\beta$ represents the number
of recursive sub-terms of each construction that we generate, whereas
$\mathcal{P}$ simply represents the probability of generating that construction
(this probability can be easily calculated based on the generation frequencies
used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}
%
The vector $\beta$ can be obtained by analyzing the shape of each source of
structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that can be calculated separately):

@@

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized.
%
In this light, \dragenp follows the same approach as \dragen, using our adapted
prediction mechanism to tune the generation probabilities of each each source of
structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
This way she can specify, for instance, a uniform distribution of data
constructors, pattern matching values and function calls or, alternative, a
distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.

\paragraph{End-to-End Prediction}
%
% So far we have shown that our prection model is able to predict the distribution
% of data constructors, composite pattern values and interface function calls that
% our automatically derived random generators will produce on average.

It is possible to provide further ``end-to-end'' guarantees on the generated
values if we restrict the generation process to only bare data constructors and
pattern matching values.
%
In particular, we can stop considering pattern matching values as atomic
constructions and start seeing them as compositions of several data
constructors.
%
Then, is possible to obtain the average ``total'' number of data constructors
that our generators will produce, regardless of if they are generated on their
own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$.

The prediction only needs to add the expected number of constructors that we
will additionally generate within pattern matching values to the prediction of
bare data constructors.
%
For instance, we can calculate the total average number of constructors |Text|
and |Join| that we will generate simply by expanding the average number of
generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
corresponding data constructors:
%
{\small
  \begin{align*}
    E^\downarrow\!\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
    E^\downarrow\!\!\left[ Join \right]
    = E\left[ Join \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
  \end{align*}
}
%
Given that each time we generate a value satisfying the first pattern matching
of the function |simplify| we add two |Text| and one |Join| data constructors to
our random value.
%
The case of the second pattern matching of |simplify| follows analogously.

Note that this expansion cannot be applied if we also generate random values
containing abstract interface function calls, as we cannot predict the output of
an arbitrary function without solving the halting problem in the process.

>>>>

\subsection{Extending predictions for structural information}

@@

Symbol $C_1 \cdots C_i$ denotes constructors, $P_1 \cdots P_j$ pattern
matchings, and $F_1 \cdots F_k$ function calls.
%
The red-light colored matrix is what we had before where the light-blue colored
cells are new---we recommend readers to get a colored copy of the paper.
%
% To extend $M$ to consider the new possible constructions, we need to calculate
% $M_{i,j}$ for every new combination of ``father'' and ``son'' constructions $i$
% and $j$, respectively.
% %

The new cells are filled as before: we need to consider the amount of holes when
generating partial pattern matching and function calls as well as their
individual probabilities.
%
% Fortunately, it is possible to follow the same pattern as we did on our original
% work, this time considering the amount of recursive pattern variables if it the
% case of a pattern matching value being a father individual, and the amount of
% recursive arguments if we deal with a abstract interface function call as a
% father individual.
%
For instance, if we consider $P_j$ as the second pattern of simplify and $F_1$
as function |div|, then the marked cell above has the value $2\cdot p_{div}$,
i.e., the amount of holes in a partially generated pattern |(Join (Join (Text s)
hole) hole)|, where |s| is some random string, times the probability to generate
a call to function |div|.
%
The rest follows analogously.

% the average amount of function calls to |div| that we will produce
% provided that we generated a value matching the second pattern of |simplify| as a
% father results $2\cdot p_{div}$.
%
% Similarly, the average amount of values satisfying the first pattern matching of
% |simplify| provided that we generated a |Tag| constructor as a father results
% $1 \cdot p_{simplify\# 1}$.

As another contribution, we found that the whole prediction process can be
factored in terms of two vectors $\beta$ and $\mathcal{P}$, such that $\beta$
represents the number of recursive sub-terms of each construction that we
generate, whereas $\mathcal{P}$ simply represents the probability of generating
that partial ADT value.%  (this probability can be easily calculated based on the
% generation frequencies used in the random generator).
%
Then, the equation \ref{eqn:prediction} can be rewriten as:
%
\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n}
\end{align*}
%
The vector $\beta$ can be obtained by analyzing the shape (namely holes) of each
source of structural information that we consider, and joining them together.
%
For instance, $\beta$ and $\mathcal{P}$ for our generation specification of HTML
values are as follows (where dashed lines indicate the composition of the
diffent sub-vectors that can be calculated separately):

@@

Note that by varying the shape of the vector $\mathcal{P}$ we can tune the
distribution of our random genrator in a way that can be always characterized
and predicted.
%
\dragenp follows a similar approach as \dragen and uses an heuristic to tune the
generation probabilities of each source of structural information.
%
This is done by runing a simulation-based optimization process at compile-time.
%
This process is parametrized by the desired distribution of values set by the
user.
%
In this manner, developers can specify, for instance, a uniform distribution of
data constructors, pattern matching values and function calls or, alternative, a
distribution of values with some constructions appearing in a different
proportion as others, e.g., two times more functions calls to |div| than |Join|
constructors.

\subsection{Overall prediction}
%\paragraph{End-to-End Prediction}
%
% So far we have shown that our prection model is able to predict the distribution
% of data constructors, composite pattern values and interface function calls that
% our automatically derived random generators will produce on average.

It is possible to provide an overall prediction of the expected number of
constructors when restricting the generation process to only bare data
constructors and pattern matching values.
%
To achieve that, we should stop considering pattern matching values as atomic
values and start seeing them as compositions of several data constructors.
%
In that manner, it is possible to obtain the expected \emph{total} number of
generated data constructors that our generators will produce---regardless if they
are generated on their own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$ and, to calculate it, we only need
to add the expected number of bare constructors that are included within each
pattern matching.
%
For instance, we can calculate the total expected number of constructors |Text|
and |Join| that we will generate by simply expanding the expected number of
generated pattern matching values $simplify\#1$ and $simplify\#2$ into their
corresponding data constructors:
%
{\small
  \begin{align*}
    E^\downarrow\!\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#2 \right] \\
    E^\downarrow\!\!\left[ Join \right]
    = E\left[ Join \right]
    + 1\!\! \cdot\!\! E\!\left[ simplify\#1 \right]
    + 2\! \cdot\!\! E\!\left[ simplify\#2 \right]
  \end{align*}
}
%
Observe that each time we generate a value satisfying the first pattern matching
of the function |simplify|, we add two |Text| and one |Join| data constructors
to our random value.
%
The case of the second pattern matching of |simplify| follows analogously.
%
Note that the overall prediction cannot be applied if we also generate random
values containing function calls, as we cannot predict the output of an abstract
function without solving the halting problem in the process.
